MS Regressions

### 2021-11-18 msdepression_regressions_started_20211118.Rmd
branched out of my big dac summary script for ease of doing things

---
title: "Dac summary 01/19/2021"
author: "Erica Baller"
date: "9/29/2021"
output:
  pdf_document: default
  html_document: default
---

title: "Dac Summary"
output:
  html_document: default
  pdf_document: default
---

@@

```{r summary}
#preprocessing - because these are sent

#load our df

#from Jan 2021 pull
data <- read.csv("/Users/eballer/BBL/msdepression/data/dac/investigatingdepressioninmspatients_dates_right_format.csv", header = TRUE, stringsAsFactors = FALSE, na.strings = c("NULL"), sep =",")

@@

#keep people with MS providers 
data_empi_acc_f_phq <- data[which(data$Provider %in% ms_providers),] #n = 17067

#keep people with ICD10 code G35 (this is the same # as people with IC9 - 340)
data_empi_acc_f_phq <- data_empi_acc_f_phq[grep("G35", data_empi_acc_f_phq$ICD10),] #n = 16830
table(data_empi_acc_f_phq$Provider)

##########################
### some preprocessing ###
##########################

#make accession # and lab findings into integer type, contained in row 28:42, Margin=2 to work on columns
data_empi_acc_f_phq$ACCESSION_NUM <- as.integer(data_empi_acc_f_phq$ACCESSION_NUM)
data_empi_acc_f_phq[27:29] <- apply(X = data_empi_acc_f_phq[27:29], FUN = as.integer, MARGIN = 2)
data_empi_acc_f_phq[31:45] <- apply(X = data_empi_acc_f_phq[31:45], FUN = as.integer, MARGIN = 2)
data_empi_acc_f_phq$sex_binarized <- ifelse(data_empi_acc_f_phq$SEX == "MALE", 1, 2)
data_empi_acc_f_phq$race_binarized <- ifelse(data_empi_acc_f_phq$RACE == "WHITE", 1, 2)
#get date into better format
data_empi_acc_f_phq$EXAM_DATE <- as.Date(data_empi_acc_f_phq$BEGIN_EXAM_DTTM, format = "%m/%d/%y")
data_empi_acc_f_phq$EXAM_DATE <- as.numeric(gsub(data_empi_acc_f_phq$EXAM_DATE, pattern = "-", replacement = ""))
########################################
#### Start counting / making subsets ###
########################################

# number of unique scanners and scans
print(paste0("Number of unique procedure names (ie. MRI Head w/out contrast; MRI Source Images): ",  length(unique(data_empi_acc_f_phq$PROC_CODE)))) #n = 10
print(paste0("Number of unique department names (ie. RAD HUP PCAM Ground): ",  length(unique(data_empi_acc_f_phq$DEPARTMENT_NAME)))) #n = 3
print(paste0("Number of unique modalities (ie. HUP PCAM GR MR4 3T; MR6 3T): ",  length(unique(data_empi_acc_f_phq$MODALITY)))) #n = 11
print(paste0("Number of unique accession #s n= ", length(unique(data_empi_acc_f_phq$ACCESSION_NUM)))) # n = 16830
print(paste0("Number of unique EMPIS n = ", length(unique(data_empi_acc_f_phq$EMPI)))) #n = 3737

#### PHQ-2 and PHQ-9 summaries

#looking at just people with phq-2 and 9
phq2_subset <- data_empi_acc_f_phq[which(!is.na(data_empi_acc_f_phq$PHQ.2)),] #n=4606

phq9_subset <- data_empi_acc_f_phq[which(!is.na(data_empi_acc_f_phq$PHQ.9)),] #n= 261

#phq-2;
print(paste0("N PHQ-2 = ", sum(!is.na(data_empi_acc_f_phq$PHQ.2)), "; n = ", length(which(phq2_subset$PHQ.2 >= 3)), " meeting criteria for depression (PHQ2) >= 3")) #n = 4606

#phq-9
print(paste0("N PHQ-9 = ", sum(!is.na(data_empi_acc_f_phq$PHQ.9)), "; noDep/mild/mod/mod-severe/severe = ", 
             length(which(phq9_subset$PHQ.9 < 5)), "/",
             length(which(phq9_subset$PHQ.9 >= 5 & phq9_subset$PHQ.9 < 10)), "/",
             length(which(phq9_subset$PHQ.9 >= 10 & phq9_subset$PHQ.9 < 15)), "/",
             length(which(phq9_subset$PHQ.9 >= 15 & phq9_subset$PHQ.9 < 20)), "/",
             length(which(phq9_subset$PHQ.9 >= 20)))) #n = 261

#both phq2 and 9

num_both_phq2_and_9 <- length(which(!is.na(data_empi_acc_f_phq$PHQ.2) & !is.na(data_empi_acc_f_phq$PHQ.9)))
print(paste0("num people with both phq2 and 9 = ", num_both_phq2_and_9)) #n = 0

#Number of unique EMPIs in phq2/phq9 subsets
unique_phq2_empis <- length(unique(phq2_subset$EMPI)) #n=966
print(paste0("N with phq2 (Unique EMPI) n = ", unique_phq2_empis, " out of ", dim(phq2_subset)[1], "; n - ", length(unique(phq2_subset$EMPI[phq2_subset$PHQ.2 >= 3])), " meeting criteria for depression (PHQ2) >= 3")) #n = 4606

unique_phq9_empis <- length(unique(phq9_subset$EMPI)) #n=76
print(paste0("N with phq9 (Unique EMPI) n = ", unique_phq9_empis, " out of ", dim(phq9_subset)[1], "; noDep/mild/mod/mod-severe/severe = ", 
             length(unique(phq9_subset$EMPI[phq9_subset$PHQ.9 < 5])), "/",
             length(unique(phq9_subset$EMPI[(phq9_subset$PHQ.9 >= 5 & phq9_subset$PHQ.9 < 10)])), "/",
             length(unique(phq9_subset$EMPI[(phq9_subset$PHQ.9 >= 10 & phq9_subset$PHQ.9 < 15)])), "/",
             length(unique(phq9_subset$EMPI[(phq9_subset$PHQ.9 >= 15 & phq9_subset$PHQ.9 < 20)])), "/",
             length(unique(phq9_subset$EMPI[phq9_subset$PHQ.9 >= 20]))))

#### ICD 9/10 for Depression
dep_df <- data_empi_acc_f_phq[grep("F3", data_empi_acc_f_phq$ICD10),] #n = 2123
num_depression <- dim(dep_df)[1]
unique_empi <- which(unique(data_empi_acc_f_phq$EMPI) %in% data_empi_acc_f_phq$EMPI) #index of the locations of unique empis; n = 3885
num_unique_EMPI_and_dep <- length(unique(dep_df$EMPI)) #n = 493

print(paste0("N with ICD 9/10 codes for depression = ", num_depression))

print(paste0("N with ICD 9/10 codes for depression and unique EMPI = ", num_unique_EMPI_and_dep, " out of ", dim(data_empi_acc_f_phq)[1]))

##### PHQ2 or 9 + ICD 9/10 Depression
num_phq2_and_dep <- length(which(!is.na(dep_df$PHQ.2))) #n = 960
print(paste0("N with ICD 9/10 codes for depression and phq2 (ALL SCANS) = ", num_phq2_and_dep, " out of ", dim(dep_df)[1]))

num_phq9_and_dep <- length(which(!is.na(dep_df$PHQ.9))) #n=140
print(paste0("N with ICD 9/10 codes for depression and phq9 (ALL SCANS) = ", num_phq9_and_dep, " out of ", dim(dep_df)[1]))

#phq-2;
print(paste0("N dep + PHQ-2 = ", sum(!is.na(dep_df$PHQ.2)), "; n = ", length(which(dep_df$PHQ.2 >= 3)), " meeting criteria for depression (PHQ2) >= 3")) #n = 4606

print(paste0("Mean dep + PHQ-2 = ", round(mean(!is.na(dep_df$PHQ.2)),3), "[",round(sd(!is.na(dep_df$PHQ.2)),3), "]; n = ", length(which(dep_df$PHQ.2 >= 3)), " meeting criteria for depression (PHQ2) >= 3")) #n = 79

#phq-9
print(paste0("N dep + PHQ-9 = ", sum(!is.na(dep_df$PHQ.9)), "; noDep/mild/mod/mod-severe/severe = ", 
             length(which(dep_df$PHQ.9 < 5)), "/",
             length(which(dep_df$PHQ.9 >= 5 & dep_df$PHQ.9 < 10)), "/",
             length(which(dep_df$PHQ.9 >= 10 & dep_df$PHQ.9 < 15)), "/",
             length(which(dep_df$PHQ.9 >= 15 & dep_df$PHQ.9 < 20)), "/",
             length(which(dep_df$PHQ.9 >= 20)))) #n = 261

print(paste0("Mean dep + PHQ-9 = ", round(mean(!is.na(dep_df$PHQ.9)),3), "[",round(sd(!is.na(dep_df$PHQ.9)),3), "]; n = ", length(which(dep_df$PHQ.9 >= 10)), " meeting criteria for depression (PHQ9) >= 10")) #n = 79

#### # unique people with dep dx and phq2/9 -> working on this
dep_dx_phq2_unique <- length(unique(dep_df$EMPI[!is.na(dep_df$PHQ.2)])) #n = 201
print(paste0("N with ICD 9/10 codes for depression and phq2 (unique) = ", dep_dx_phq2_unique, " out of ", dim(dep_df)[1]))

dep_dx_phq9_unique <- length(unique(dep_df$EMPI[!is.na(dep_df$PHQ.9)])) #n = 43
print(paste0("N with ICD 9/10 codes for depression and phq9 (unique) = ", dep_dx_phq9_unique, " out of ", dim(dep_df)[1]))

####

#### ICD 9/10 for Healthy Group
healthy_df <- data_empi_acc_f_phq[(-grep("F3", data_empi_acc_f_phq$ICD10)),] #n = 14707
num_healthy <- dim(healthy_df)[1]
unique_empi <- which(unique(healthy_df$EMPI) %in% healthy_df$EMPI) #index of the locations of unique empis; n = 3244
num_unique_EMPI_and_healthy <- length(unique(healthy_df$EMPI)) #n = 3244

print(paste0("N without Depression/ICD 9/10 codes for healthy = ", num_healthy))

print(paste0("N without Depression/ ICD 9/10 codes for healthy and unique EMPI = ", num_unique_EMPI_and_healthy, " out of ", dim(healthy_df)[1]))

##### PHQ2 or 9 + No Depression ICD9/10
num_phq2_and_healthy <- length(which(!is.na(healthy_df$PHQ.2))) #n = 3646
print(paste0("N without Depression/ ICD 9/10 codes for healthy and phq2 (ALL SCANS) = ", num_phq2_and_healthy, " out of ", dim(healthy_df)[1]))

#phq-2;
print(paste0("N healthy w/PHQ-2 = ", sum(!is.na(healthy_df$PHQ.2)), "; n = ", length(which(healthy_df$PHQ.2 >= 3)), " meeting criteria for depression (PHQ2) >= 3")) #n = 47

print(paste0("Mean/SD healthy w/PHQ-2 = ", round(mean(!is.na(healthy_df$PHQ.2)),3), "[",round(sd(!is.na(healthy_df$PHQ.2)),3), "]; n = ", length(which(healthy_df$PHQ.2>= 3)), " meeting criteria for depression (PHQ2) >= 3"))

num_phq9_and_healthy <- length(which(!is.na(healthy_df$PHQ.9))) #n=121
print(paste0("N without Depression/ ICD 9/10 codes for healthy and phq9 (ALL SCANS) = ", num_phq9_and_healthy, " out of ", dim(healthy_df)[1]))

#phq-9
print(paste0("N healthy w/PHQ-9 = ", sum(!is.na(healthy_df$PHQ.9)), "; noDep/mild/mod/mod-severe/severe = ", 
             length(which(healthy_df$PHQ.9 < 5)), "/",
             length(which(healthy_df$PHQ.9 >= 5 & healthy_df$PHQ.9 < 10)), "/",
             length(which(healthy_df$PHQ.9 >= 10 & healthy_df$PHQ.9 < 15)), "/",
             length(which(healthy_df$PHQ.9 >= 15 & healthy_df$PHQ.9 < 20)), "/",
             length(which(healthy_df$PHQ.9 >= 20)))) #n = 55/34/21/1/10

print(paste0("Mean/SD healthy w/PHQ-9 = ", round(mean(!is.na(healthy_df$PHQ.9)),3), "[",round(sd(!is.na(healthy_df$PHQ.9)),3), "]; n = ", length(which(healthy_df$PHQ.9>= 10)), " meeting criteria for depression (PHQ9) >= 10"))

#statistical testing
t.test(!is.na(healthy_df$PHQ.2), !is.na(dep_df$PHQ.2))
t.test(!is.na(healthy_df$PHQ.9), !is.na(dep_df$PHQ.9))

#test first instance of phq2/9
df_unique_empi_healthy <- healthy_df %>% 
 group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup()

df_unique_empi_dep <- dep_df %>% 
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup()

t.test(!is.na(df_unique_empi_healthy$PHQ.2), !is.na(df_unique_empi_dep$PHQ.2))
t.test(!is.na(df_unique_empi_healthy$PHQ.9), !is.na(df_unique_empi_dep$PHQ.9))

print(paste0("Mean/SD depressed w/PHQ-2/9; healthy w/ph2/9 = ",
             round(mean(!is.na(df_unique_empi_dep$PHQ.2)),3),"[",round(sd(!is.na(df_unique_empi_dep$PHQ.2)),3), "];", 
             round(mean(!is.na(df_unique_empi_dep$PHQ.9)),3),
              "[",round(sd(!is.na(df_unique_empi_dep$PHQ.9)),3), "];",
             round(mean(!is.na(df_unique_empi_healthy$PHQ.2)),3),
             "[",round(sd(!is.na(df_unique_empi_healthy$PHQ.2)),3), "];",
             round(mean(!is.na(df_unique_empi_healthy$PHQ.9)),3),
             "[",round(sd(!is.na(df_unique_empi_healthy$PHQ.9)),3), "]"))

#### # unique people without dep dx and phq2/9 
dep_dx_phq2_unique <- length(unique(healthy_df$EMPI[!is.na(healthy_df$PHQ.2)])) #n = 201
print(paste0("N without Depression/ ICD 9/10 codes for healthy and phq2 (unique) = ", dep_dx_phq2_unique, " out of ", dim(healthy_df)[1]))

dep_dx_phq9_unique <- length(unique(healthy_df$EMPI[!is.na(healthy_df$PHQ.9)])) #n = 43
print(paste0("N without Depression/ ICD 9/10 codes for healthy and phq9 (unique) = ", dep_dx_phq9_unique, " out of ", dim(healthy_df)[1]))

####

#demographics and lab historgrams
#print histogram of phq-2/9s for people

hist(data_empi_acc_f_phq$PHQ.2[which(!is.na(data_empi_acc_f_phq$PHQ.2))], main = paste0("PHQ-2 Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$PHQ.2)))), xlab = "PHQ-2", breaks = 8)

hist(data_empi_acc_f_phq$PHQ.9[which(!is.na(data_empi_acc_f_phq$PHQ.9))], main = paste0("PHQ-9 Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$PHQ.9)))), xlab = "PHQ-9", breaks = 30)

#fcode histos
hist(dep_df$PHQ.2, main = paste0("PHQ2 and Dep Histo : n = ", num_phq2_and_dep), xlab = paste0("PHQ2"))

hist(dep_df$PHQ.9,main = paste0("PHQ9 and Dep Histo : n = ", num_phq9_and_dep), xlab = paste0("PHQ9"))

###

#age histogram
hist(data_empi_acc_f_phq$MRI_ENC_AGE[which(!is.na(data_empi_acc_f_phq$MRI_ENC_AGE))], main = paste0("Age Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$MRI_ENC_AGE)))), xlab = "AGE AT MRI ENCOUNTER")

#tsh
hist(data_empi_acc_f_phq$TSH[which(!is.na(data_empi_acc_f_phq$TSH))], main = paste0("TSH Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$TSH)))), xlab = "TSH")

#hemoglobin
hist(data_empi_acc_f_phq$hemoglobin[which(!is.na(data_empi_acc_f_phq$hemoglobin))], main = paste0("Hemoglobin Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$hemoglobin)))), xlab = "hemoglobin")

#wbc
hist(data_empi_acc_f_phq$WBC[which(!is.na(data_empi_acc_f_phq$WBC))], main = paste0("WBC Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$WBC)))), xlab = "WBC")

#folate
hist(data_empi_acc_f_phq$FOLATE[which(!is.na(data_empi_acc_f_phq$FOLATE))], main = paste0("Folate Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$FOLATE)))), xlab = "folate", breaks = 10)

#vitamin D
hist(data_empi_acc_f_phq$VitD[which(!is.na(data_empi_acc_f_phq$VitD))], main = paste0("Vitamin D Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$VitD)))), xlab = "Vitamin D", breaks = 10)

#race
hist(data_empi_acc_f_phq$race_binarized[which(!is.na(data_empi_acc_f_phq$race_binarized))], 
     main = paste0("Race Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$race))), "; White = ", length(which(data_empi_acc_f_phq$race_binarized == 1)), "; NW = ", length(which(data_empi_acc_f_phq$race_binarized == 2))),
     xlab = "white = 1, other = 2", 
     breaks = 2)

#f-codes - people usually have BOTH icd9/10 or neither

#Gender breakdown
#sex
hist(data_empi_acc_f_phq$sex_binarized[which(!is.na(data_empi_acc_f_phq$sex_binarized))], 
     main = paste0("Sex Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$sex_binarized))), "; M = ", length(which(data_empi_acc_f_phq$sex_binarized == 1)), "; F = ", length(which(data_empi_acc_f_phq$sex_binarized == 2)) ), 
     xlab = "Male = 1, Female = 2", 
     breaks = 2)

#sex with depression diagnosis
hist(dep_df$sex_binarized[which(!is.na(dep_df$sex_binarized))], 
     main = paste0("Sex Histo in ICD9/10 depression dx group : n = ", length(which(!is.na(dep_df$sex_binarized))), "; M = ", length(which(dep_df$sex_binarized == 1)), "; F = ", length(which(dep_df$sex_binarized == 2)) ), 
     xlab = "Male = 1, Female = 2", 
     breaks = 2)

#sex w/ phq2
hist(phq2_subset$sex_binarized[which(!is.na(phq2_subset$sex_binarized))], 
     main = paste0("Sex Histo w/phq2: n = ", length(which(!is.na(phq2_subset$sex_binarized))), "; M = ", length(which(phq2_subset$sex_binarized == 1)), "; F = ", length(which(phq2_subset$sex_binarized == 2)) ), 
     xlab = "Male = 1, Female = 2", 
     breaks = 2)

#sex w/phq9
hist(phq9_subset$sex_binarized[which(!is.na(phq9_subset$sex_binarized))], 
     main = paste0("Sex Histo w/phq9: n = ", length(which(!is.na(phq9_subset$sex_binarized))), "; M = ", length(which(phq9_subset$sex_binarized == 1)), "; F = ", length(which(phq9_subset$sex_binarized == 2)) ), 
     xlab = "Male = 1, Female = 2", 
     breaks = 2)

```{r frequencies_of_dep_in_phq2_vs_9}
num_dep_phq2 <- length(which(!is.na(data_empi_acc_f_phq$PHQ.2) & (data_empi_acc_f_phq$PHQ.2 >= 3))) #n = 130
num_dep_phq9 <- length(which(!is.na(data_empi_acc_f_phq$PHQ.9) & (data_empi_acc_f_phq$PHQ.9 >= 5))) #n = 184;
#https://www.sciencedirect.com/science/article/pii/B9780128179338000293 recommends using cut point of 11
num_total_phq2 <-length(which(!is.na(data_empi_acc_f_phq$PHQ.2)))
num_total_phq9 <-length(which(!is.na(data_empi_acc_f_phq$PHQ.9))) 
dep_total_phq_table <- data.frame(matrix(data = c(num_dep_phq2, num_dep_phq9, num_total_phq2, num_total_phq9), nrow = 2, ncol = 2))
names(dep_total_phq_table) <- c("dep", "healthy")
row.names(dep_total_phq_table) <- c("phq2", "phq9")
chisq_dep_phq2_v_phq9 <- chisq.test(dep_total_phq_table)
chisq_dep_phq2_v_phq9
print("observed")
chisq_dep_phq2_v_phq9$observed
print("expected")
chisq_dep_phq2_v_phq9$expected
print("residuals")
chisq_dep_phq2_v_phq9$residuals

```
```{r chi_sq_dep_versus_healthy_PHQ2_and_9}
num_dep_phq2 <- length(which(!is.na(dep_df$PHQ.2) & (dep_df$PHQ.2 >= 3))) #n = 83
num_dep_phq9 <- length(which(!is.na(dep_df$PHQ.9) & (dep_df$PHQ.9 >= 10))) #n = 79
num_healthy_phq2 <- length(which(!is.na(healthy_df$PHQ.2) & (healthy_df$PHQ.2 >= 3))) #n = 47
num_healthy_phq9 <- length(which(!is.na(healthy_df$PHQ.9) & (healthy_df$PHQ.9 >= 10))) #n = 32;
total_dep_phq2 <- length(which(!is.na(dep_df$PHQ.2))) #n = 83
total_dep_phq9 <- length(which(!is.na(dep_df$PHQ.9))) #n = 79
total_healthy_phq2 <- length(which(!is.na(healthy_df$PHQ.2))) #n = 47
total_healthy_phq9 <- length(which(!is.na(healthy_df$PHQ.9))) #n = 32;

#### phq2 symptomatic versus asymptomatic
phq2_table <- data.frame(matrix(data = c(num_dep_phq2,(total_dep_phq2-num_dep_phq2), num_healthy_phq2,   (total_healthy_phq2-num_healthy_phq2)), nrow = 2, ncol = 2))
names(phq2_table) <- c("dep", "healthy")
row.names(phq2_table) <- c("symptomatic", "asymptomatic")
chisq_dep_phq2 <- chisq.test(phq2_table)
chisq_dep_phq2
print("observed")
chisq_dep_phq2$observed
print("expected")
chisq_dep_phq2$expected
print("residuals")
chisq_dep_phq2$residuals

#### phq9 symptomatic versus asymptomatic
phq9_table <- data.frame(matrix(data = c(num_dep_phq9,(total_dep_phq9-num_dep_phq9), num_healthy_phq9,   (total_healthy_phq9 - num_healthy_phq9)), nrow = 2, ncol = 2))
names(phq9_table) <- c("dep", "healthy")
row.names(phq9_table) <- c("symptomatic", "asymptomatic")
chisq_dep_phq9 <- chisq.test(phq9_table)
chisq_dep_phq9
print("observed")
chisq_dep_phq9$observed
print("expected")
chisq_dep_phq9$expected
print("residuals")
chisq_dep_phq9$residuals

################### Unique

num_dep_phq2 <- length(which(!is.na(dep_df$PHQ.2) & (dep_df$PHQ.2 >= 3))) #n = 83
num_dep_phq9 <- length(which(!is.na(dep_df$PHQ.9) & (dep_df$PHQ.9 >= 10))) #n = 79
num_healthy_phq2 <- length(which(!is.na(healthy_df$PHQ.2) & (healthy_df$PHQ.2 >= 3))) #n = 47
num_healthy_phq9 <- length(which(!is.na(healthy_df$PHQ.9) & (healthy_df$PHQ.9 >= 10))) #n = 32;
total_dep_phq2 <- length(which(!is.na(dep_df$PHQ.2))) #n = 83
total_dep_phq9 <- length(which(!is.na(dep_df$PHQ.9))) #n = 79
total_healthy_phq2 <- length(which(!is.na(healthy_df$PHQ.2))) #n = 47
total_healthy_phq9 <- length(which(!is.na(healthy_df$PHQ.9))) #n = 32;

#patients who are depressed who score on phq2/9
num_dep_phq2_unique <- length(unique(dep_df$EMPI[which(dep_df$PHQ.2 >= 3)]))
num_dep_phq9_unique <- length(unique(dep_df$EMPI[which(dep_df$PHQ.9 >= 10)]))

#persons who are characterized as healthy but score on PHQ2/9
num_healthy_phq2_unique <- length(unique(healthy_df$EMPI[which(healthy_df$PHQ.2 >= 3)]))
num_healthy_phq9_unique <- length(unique(healthy_df$EMPI[which(healthy_df$PHQ.9 >= 10)]))

#total # of people who have phq2/9
asym_dep_phq2_unique <- length(unique(dep_df$EMPI[which(!is.na(dep_df$PHQ.2))])) - num_dep_phq2_unique
asym_dep_phq9_unique <- length(unique(dep_df$EMPI[which(!is.na(dep_df$PHQ.9))])) - num_dep_phq9_unique
asym_healthy_phq2_unique <- length(unique(healthy_df$EMPI[which(!is.na(healthy_df$PHQ.2))])) - num_healthy_phq2_unique
asym_healthy_phq9_unique <- length(unique(healthy_df$EMPI[which(!is.na(healthy_df$PHQ.9))])) - num_healthy_phq9_unique

#### phq2 symptomatic versus asymptomatic
phq2_table_unique <- data.frame(matrix(data = c(num_dep_phq2_unique,asym_dep_phq2_unique, num_healthy_phq2_unique, asym_healthy_phq2_unique), nrow = 2, ncol = 2))
names(phq2_table_unique) <- c("dep", "healthy")
row.names(phq2_table_unique) <- c("symptomatic", "asymptomatic")
chisq_dep_phq2_unique <- chisq.test(phq2_table_unique)
chisq_dep_phq2_unique
print("observed")
chisq_dep_phq2_unique$observed
print("expected")
chisq_dep_phq2_unique$expected
print("residuals")
chisq_dep_phq2_unique$residuals

#### phq9 symptomatic versus asymptomatic
phq9_table_unique <- data.frame(matrix(data = c(num_dep_phq9_unique,asym_dep_phq9_unique, num_healthy_phq9_unique,   asym_healthy_phq9_unique), nrow = 2, ncol = 2))
names(phq9_table_unique) <- c("dep", "healthy")
row.names(phq9_table_unique) <- c("symptomatic", "asymptomatic")
chisq_dep_phq9_unique <- chisq.test(phq9_table_unique)
chisq_dep_phq9_unique
print("observed")
chisq_dep_phq9_unique$observed
print("expected")
chisq_dep_phq9_unique$expected
print("residuals")
chisq_dep_phq9_unique$residuals

This next step, put together 9/29/2021, aims to define our control and case groups based on the following definitions-

Healthy group - No F33* code, and PHQ2 and 9 always 0

Lifetime depression - EITHER F33* diagnosis OR PHQ2 >= 3 OR PHQ9 >= 10
   - of note:
    *PHQ9 0-4 - no depression
    * 5-9 - mild depression
    * 10-14 - moderate depression
    * 15-19 - mod/severe
    * 20+ - severe

Logic
  1) First, separate out people with ICD10 code (F3*) into depressed and non depressed
       - dep_df is people with an F3* code
       - healthy df is people without
       - look for unique EMPIs in each of these groups

  2) Within "healthy" group, separate who has EVER had a lifetime history of testing positive on PHQ2 
        - Depression is PHQ2 >=3 at any point
        - Healthy if PHQ2 = 0
        - Look for unique EMPIs in these groups

  3) Within "healthy" group, separate who has EVER had a lifetime history of testing positive on PHQ9
        - Depression is PHQ9 >= 10 at any point
        - Healthy if PHQ9 = 0
        - Look for unique EMPIs in these groups

  4) Combine for depressed group and healthy
        - Look for people with PHQ2 or 9 meeting criteria for depression that aren't in depression group, and add them -> dep_num_icd10_plus_phq2_and_phq9
        - remove people from healthy_df if people had EMPIs in phq2 or 9 depressed group -> healthy_num_no_icd10_plus_phq2_and_phq9
        - make separate group that only includes healthies for whom we have phq2 or 9 -> healthy_num_no_icd10_plus_phq2_and_phq9_MUST_HAVE_PHQ2_OR_9

```{r defining_control_and_dep_groups}

##########################################################
#### Step 1- make F3* depressed and healthy groups* ######
##########################################################

########
### separate out depressed group
########
print("separate out depressed group")
dep_df <- data_empi_acc_f_phq[grep("F3", data_empi_acc_f_phq$ICD10),] #n = 2123
num_depression <- dim(dep_df)[1]
num_unique_EMPI_and_dep <- length(unique(dep_df$EMPI)) #n = 493
print(paste0("N with ICD 9/10 codes for depression = ", num_depression))
print(paste0("N with ICD 9/10 codes for depression and unique EMPI in whole depression group = ", num_unique_EMPI_and_dep, " out of ", num_depression))

#######
### separate out healthy group
#######
print("separate out healthy group")
healthy_df <- data_empi_acc_f_phq[-grep("F3", data_empi_acc_f_phq$ICD10),] #n = 14707
num_healthy <- dim(healthy_df)[1] 
num_unique_EMPI_and_healthy <- length(unique(healthy_df$EMPI)) #n = 3244
print(paste0("N with WITHOUT depression ICD 9/10 codes (i.e. \"healthy\"): ", num_healthy))
print(paste0("N with WITHOUT ICD 9/10 depression codes for depression and unique EMPI = ", num_unique_EMPI_and_healthy, " out of ", num_healthy))

@@

print("Get people in healthy group with phq2s for dep/healthy")
# num in healthy group with PHQ2
phq2_subset <- healthy_df[which(!is.na(healthy_df$PHQ.2)),] #n=3646, unique 698
depressed_phq2_subset <- phq2_subset[which(phq2_subset$PHQ.2 >= 3),] #n = 47 
depressed_unique_phq2_empis <- unique(depressed_phq2_subset$EMPI) #n=10
healthy_phq2_subset <- phq2_subset[which(phq2_subset$PHQ.2 == 0),] #n = 3326
healthy_unique_phq2_empis <- unique(healthy_phq2_subset$EMPI) #n = 631

print(paste0("PHQ2 breakdown in healthy group (total n with score = ", dim(phq2_subset)[1], "[unique = ", length(unique(phq2_subset$EMPI)),"]): PHQ2 >= 3 :", dim(depressed_phq2_subset)[1], "[unique = ", length(depressed_unique_phq2_empis), "]; PHQ2 == 0 : ", dim(healthy_phq2_subset)[1], "[unique = ", length(healthy_unique_phq2_empis),"]"))

@@

@@

```{r dep_vs_healthy_regressions}

>>>>

---
title: "MS Regressions"
author: "Erica Baller"
date: "11/18/2021"
output:
  html_document: default
  pdf_document: default
---

@@

homedir <- "/Users/eballer/BBL/msdepression/"

#from Jan 2021 pull
data <- read.csv(paste0(homedir,"/data/dac/investigatingdepressioninmspatients_dates_right_format.csv"), header = TRUE, stringsAsFactors = FALSE, na.strings = c("NULL"), sep =",")

@@

columns_to_make_integer <- c("ACCESSION_NUM", "PAT_AGE_AT_EXAM","MRI_ENC_AGE","CURRENT_AGE","hemoglobin", "WBC","RBC","B12","FOLATE", "TSH", "RPR","CSFAPPEAR1","CSFAPPEAR4","CSFCOLOR1",       "CSFCOLOR4","CSFTUBE","CSFTUBE4","VitD","PHQ.2","PHQ.9")

##########################
### some preprocessing ###
##########################
#we need to keep scans from people seen by an MS provider (n = 17067) who have an MS diagnostic code (n=16,830)
#we make a bunch of columns from character to integer, binarize sex and race, and put date into a nice format YYYYMMDD, the %m%d%y indicates what it started out as

#goal is to keep people who were seen by 
 data_empi_acc_f_phq <- data %>% 
   filter(Provider %in% ms_providers) %>%       #n=17,067
   filter(grepl("G35", ICD10)) %>%              #n=16830
   mutate(across(.cols = columns_to_make_integer, .fns = as.integer)) %>%
   mutate(sex_binarized = ifelse(SEX == "MALE", 1, 2)) %>%
   mutate(osex = ordered(sex_binarized,levels = c(1,2), labels = c("Male","Female"))) %>%
   mutate(race_binarized = ifelse(RACE == "WHITE", 1, 2)) %>%
   mutate(orace = ordered(race_binarized,levels = c(1,2), labels = c("White","Non-white"))) %>%
   mutate(EXAM_DATE = as.Date(BEGIN_EXAM_DTTM, format = "%m/%d/%y")) %>% 
   mutate(EXAM_DATE = gsub(EXAM_DATE, pattern = "-", replacement = ""))

########
### separate out depressed group
########
print("separate out depressed group")
dep_df <- data_empi_acc_f_phq %>% filter(grepl("F3", ICD10)) #n = 2123
num_depression <- dim(dep_df)[1]
num_unique_EMPI_and_dep <- length(unique(dep_df$EMPI)) #n = 493
print(paste0("N with ICD 9/10 codes for depression = ", num_depression))
print(paste0("N with ICD 9/10 codes for depression and unique EMPI in whole depression group = ", num_unique_EMPI_and_dep, " out of ", num_depression))

#find number of participants who meet criteria for depression in the PHQ2/9 symptom rating
print(paste0("N with ICD 9/10 codes for depression, PHQ2 >= 3, unique = ", length(!is.na(unique(dep_df$EMPI[dep_df$PHQ.2 >= 3]))))) #n=23
print(paste0("N with ICD 9/10 codes for depression, PHQ9 >= 10, unique = ", length(!is.na(unique(dep_df$EMPI[dep_df$PHQ.9 >= 10])))))#n=26

#######
### separate out healthy group
#######
print("separate out healthy group")
healthy_df <- data_empi_acc_f_phq %>% filter(!grepl("F3", ICD10)) #n = 14707
num_healthy <- dim(healthy_df)[1] 
num_unique_EMPI_and_healthy <- length(unique(healthy_df$EMPI)) #n = 3244
print(paste0("N with WITHOUT depression ICD 9/10 codes (i.e. \"healthy\"): ", num_healthy))
print(paste0("N with WITHOUT ICD 9/10 depression codes for depression and unique EMPI = ", num_unique_EMPI_and_healthy, " out of ", num_healthy))

@@

print("Get people in healthy group with phq2s for dep/healthy")
# num in healthy group with PHQ2
phq2_subset <- healthy_df %>% filter(!is.na(PHQ.2)) #n=3646, unique 698
depressed_phq2_subset <- phq2_subset %>% filter(PHQ.2 >= 3) #n = 47 
depressed_unique_phq2_empis <- unique(depressed_phq2_subset$EMPI) #n=10
healthy_phq2_subset <- phq2_subset %>% filter(PHQ.2 == 0) #n = 3326
healthy_unique_phq2_empis <- unique(healthy_phq2_subset$EMPI) #n = 631

print(paste0("PHQ2 breakdown in healthy group (total n with score = ", dim(phq2_subset)[1], "[unique = ", length(unique(phq2_subset$EMPI)),"]): PHQ2 >= 3 :", dim(depressed_phq2_subset)[1], "[unique = ", length(depressed_unique_phq2_empis), "]; PHQ2 == 0 : ", dim(healthy_phq2_subset)[1], "[unique = ", length(healthy_unique_phq2_empis),"]")) #PHQ2 breakdown in healthy group (total n with score = 3646[unique = 698]): PHQ2 >= 3 :47[unique = 10]; PHQ2 == 0 : 3326[unique = 631]

@@

write.table(fascicle_names,"/Users/eballer/BBL/msdepression/templates/dti/HCP_YA1065_tractography/fascicle_names.csv", row.names = F, quote = F, col.names = F)

@@

```{r dep_vs_healthy_regressions}
##########################################################
#                Step 9 - Regressions                    #
##########################################################
